---
title: "Final Project"
author: "Sai Charan Chintala, Nidhi Vraj Sadhuvala"
date: "4/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ade4)
library(ggrepel)
library(ggbiplot)
library(reshape)
library(dplyr)
library(viridis)
library(devtools)
library(MASS)
library(tidyverse)
library(broom)
library(plyr)
library(ggthemes)
```
## R Markdown

```{r}
data <- data.frame(read.csv("/Users/nidhisaduvala/Downloads/imdb_processed.csv", header = TRUE))
```

```{r}
colSums(is.na(data))
```

```{r}
data1<-data[complete.cases(data),]
colSums(is.na(data1))
data3<- data1[data1$kind=="movie",]
data3<- data3[data3$runtime<quantile(data1$runtime,0.97),]
data1<-data3
```

```{r}
ggplot(data=data1,aes(x = rating))+ ylab("number of movies") +
  geom_histogram(binwidth=0.5) + ggtitle("Histogram plot of movie ratings") + geom_vline(aes(xintercept = median(rating)), colour="red", size=1, linetype="dashed") + geom_vline(aes(xintercept = mean(data1$rating)), colour="blue", size=1, linetype="dashed")+ annotate("text", x = mean(data1$rating) * 1.7, y = mean(data1$rating) * 1.7, label = paste("Mean =", round(mean(data1$rating),1)), col = "blue",size = 4,hjust=0.8,vjust=-1) + annotate("text", x = median(data1$rating) * 1.7, y = median(data1$rating) * 1.7, label = paste("Median =", median(data1$rating)), col = "red",size = 4,hjust=1,vjust=1)
```


```{r}
ggplot(data=data1,aes(x = runtime))+
  geom_histogram(binwidth=5)+
  xlim(c(0,quantile(data1$runtime,0.99)))+
  geom_vline(aes(xintercept = median(runtime)), colour="red", size=1, linetype="dashed") + geom_vline(aes(xintercept = mean(data1$runtime)), colour="blue", size=1, linetype="dashed")+ annotate("text", x = mean(data1$runtime) * 1.7, y = mean(data1$runtime) * 1.7, label = paste("Mean =", round(mean(data1$runtime),1)), col = "blue",size = 4,hjust=0,vjust=-1) + annotate("text", x = median(data1$runtime) * 1.7, y = median(data1$runtime) * 1.7, label = paste("Median =", median(data1$runtime)), col = "red",size = 4,hjust= 0,vjust=1) + ggtitle("Runtime Histogram") + ylab('count')

```

```{r}
ggplot(data=data1,aes(x = year))+
  geom_histogram(binwidth=1)+geom_vline(xintercept=1945,col="blue")+
    annotate("text",x=1945+5,y=4.5,label="WW II",vjust=-3,
        col="blue",fontface="bold")+
    geom_vline(xintercept=1934,col="darkgreen",lwd=12,alpha=1/2)+
    annotate("text",x=1934-8,y=4.5,vjust=-3,
        label="Great\nDepression",col="darkgreen",fontface="bold")+ylab("count of movies") + ggtitle("Number of movies released across all the years")


```
1. movies count in great depression is less compared to before and after great depression
2. World war 2 won't have much affect on movies count.
3. Count of movies is increasing exponentiallu over the years

```{r}
ggplot(data=data1,aes(x = runtime))+
  geom_histogram(binwidth=3)+
  xlim(c(0,quantile(data1$runtime,0.99))) + ggtitle("Runtime Histogram")
```
1. Most of the movies have runtime between 90 and 120 minutes.
2. Slightly skewed to right

```{r}
ggplot(data=data1,aes(x = vote))+
  geom_histogram(binwidth=500)+
  xlim(c(0,quantile(data1$vote,0.9))) + ggtitle("Votes histogram")
```
1. Here the bin size is 500. Most of the movies have votes between 0-500 and decreasing gradually.
2. Distribution is highly skewed to right

```{r}
ggplot(data=data1,aes(x=year,y=rating))+
  geom_point(alpha=1/8,position="jitter",scale=0.8)+ 
  geom_line(stat="summary",fun.y=median,col=2,linetype=1,size =1)+geom_vline(xintercept=1945,col="blue")+
    annotate("text",x=1945+5,y=4.5,label="WW II",
        col="blue",fontface="bold")+
    geom_vline(xintercept=1934,col="darkgreen",lwd=12,alpha=1/2)+
    annotate("text",x=1934-8,y=4.5,
        label="Great\nDepression",col="darkgreen",fontface="bold")+ ggtitle("Average Ratings and movies count over the years")
```
1.Medians across the years are almost stable 1960 and gradually movieing decreasing trend
2. There is no significant decrease in ratings indicates that there is no quality loss during great depression and ww2

```{r}
ggplot(data1, aes(x = vote, y = rating)) +
          geom_point()+ggtitle("Average Rating vs Votes")
```
1. More than 90% of the movies have votes between 0-500k
2. The movie with the highest number of votes is the 'the Shawshank Redemption" with average rate above 9.3


1. Most of the movies have duration between 1.5-2hrs. 
2. And most of these movies have ratings between 6 and 8.
3. However, as the duration increases, distribution of the ratings decreases.

```{r}
ggplot(data1, aes(x = year, y = rating)) +
          geom_bin2d() +
          geom_smooth(color="black") +
          scale_x_continuous() +
          scale_y_continuous(breaks = 1:10) +
          scale_fill_viridis_c(option = "plasma", labels = comma, trans = 'log10')+geom_vline(xintercept=1945,col=2)+
    annotate("text",x=1945+5,y=4.5,label="WW II",
        col=2,fontface="bold")+
    geom_vline(xintercept=1934,col="darkgreen",lwd=12,alpha=1/2)+
    annotate("text",x=1934-8,y=4.5,
        label="Great\nDepression",col="darkgreen",fontface="bold")+ ggtitle("Linear model and relationship between ratings and count over the years")
```
1. Distribution of ratings is increased over time.
2. Linearly fitted ratings lines have decreased over time.
3. Movies are increasing over the years
4. Many movies have ratings between 6-8 in most of the years

```{r}
ggplot(data1, aes(x = vote, y = rating)) +
          geom_bin2d() +
          geom_smooth(color = "red") +
          scale_x_log10(labels = comma) +
          scale_y_continuous(breaks = 1:10) +
          scale_fill_viridis_c(labels = comma) + ggtitle("Votes vs rating heatmap")
```
1. More movies have votes between 1000 and 10000 with average rating between 6-8
2. As the number of votes increasing, distribution of ratings going towards higher values

```{r}
data1<-data[complete.cases(data),]
colSums(is.na(data1))
data3<- data1[data1$kind=="movie",]
data3<- data3[data3$runtime<quantile(data1$runtime,0.97),]
data1<-data3
```

```{r}
data3<- data1[data1$kind=="movie",]
data3<- data3[data3$runtime<quantile(data1$runtime,0.97),]
```

```{r}
data3[data3==''] <- NA
data3 = data3[!is.na(data3$genre),]
data3 = data3[!is.na(data3$director),]
data3['genres_sep'] = gsub("\\[|\\]", "", data3$genre)
data3['director_sep'] = gsub("\\[|\\]", "", data3$director)
```

```{r}
for (i in 1:length(data3$genre)){
  data3[i,'primary_genre'] = list(strsplit(data3$genres_sep, ",")[[i]][1])
}
```

```{r}
data3$primary_genre<-gsub("'","",data3$primary_genre)
```

```{r}
for (i in 1:length(data3$director)){
  data3[i,'primary_director'] = list(strsplit(data3$director_sep, ",")[[i]][1])
}
```

```{r}
data3$primary_director<-gsub("'","",data3$primary_director)
```

#Does the movie genres have an impact on ratings and votes?
```{r}
ratingdat <- ddply(data3, c("primary_genre"), summarise,
                   M = mean(rating, na.rm = TRUE),
                   SE = sd(rating)/sqrt(length(rating)),
                   N = length(rating))
ratings<-ratingdat[which(ratingdat$N>=100),]
```

```{r}




ratings$primary_genre <- factor(ratings$primary_genre)
ratings$primary_genre <- reorder(ratings$primary_genre, ratings$M)

ggplot(ratings,aes(x=M,xmin=M-SE,xmax=M+SE,y=primary_genre))+
  geom_point()+
  geom_segment(aes(x=M-SE,xend=M+SE,y= primary_genre,yend=primary_genre))+
  theme(axis.text = element_text(size=8))+xlab("Mean Rating")+ylab("Genre")

```


```{r}

data3[data3==''] <- NA
data3 = data3[!is.na(data3$language),]
data3['language_sep'] = gsub("\\[|\\]", "", data3$language)

for (i in 1:length(data3$language)){
  data3[i,'primary_language'] = list(strsplit(data3$language_sep, ",")[[i]][1])
}

data3$primary_language<-gsub("'","",data3$primary_language)


ratingdat_lan <- ddply(data3, c("primary_language"), summarise,
                   M = mean(rating, na.rm = TRUE),
                   SE = sd(rating)/sqrt(length(rating)),
                   N = length(rating))
ratings_lan<-ratingdat_lan[which(ratingdat_lan$N>=200),]

```


```{r}
primary_three_language=data3[data3$primary_language == c("English", "Japanese", "French"),]
```


```{r}
ggplot(data = primary_three_language) + 
  aes(x = primary_language, y = rating, fill = primary_language) + 
  geom_boxplot()
```


#top director
```{r}
directors_grp = data3 %>% group_by(primary_director)  %>%
                    summarise(avg_ratings = mean(rating),
                              avg_votes = mean(vote),
                              success_rate = (avg_ratings*avg_votes),
                              list_titles = list(title),
                              .groups = 'drop')
```


```{r}
top_director <- directors_grp[order(-directors_grp$success_rate),]
head(top_director)
```

```{r}
top_director$list_titles[] <- lapply(top_director$list_titles, as.character)

```

```{r}
top_director %>%
  dplyr::select(primary_director,list_titles)
```

```{r}
top5_rating <- top_director[order(top_director$success_rate,decreasing = T),][1:5,]
top5_rating
```

```{r}
head(top5_rating$list_titles)
```

